Stack Safety as a Security Property

### 2020-07-09 main.tex
wip on 3

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  What is stack safety? While many mechanisms have been proposed to
  protect against attacks that influence the call stack of a program,
  and many others rely on a protected stack as an assumption to
  provide higher level guarantees, a clear stack safety definition
  is missing from the literature.
%
  In this paper we give such a formal definition of stack safety as a
  security property.
  %
  We also explore how this formal definition can be enforced by
  previously proposed security mechanisms and discuss provably
  stronger and weaker variations of interest.
\end{abstract}

@@

\apt{Need to clarify to what extent our properties depend on the
particular compilation style, calling sequences, etc. illustrated here.}

\section{Threat Model}
\label{sec:threat}

@@

Stack safety is meant to protect a caller from a confused or malicious callee's
attempts to access private data or highjack its control flow. Our
attacker model reflects this intention by assuming that the attacker
can run arbitrary code to try to access or overwrite its caller's stack data
or disrupt its control state. Hardware and timing attacks are out of
scope. \leo{Anything else to add here?}
\leo{Maybe a good place to discuss calling conventions and that they
  are essentially enforced by the policy or sth?}
\rb{Intuition that the callee can do anything in principle, but the machine will
halt if it does not follow the calling convention, etc.?}

\apt{Informal definition of policies as auxiliary state here.}

Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
policy state to the ordinary machine state\bcp{This may be surprising to
  people who are not thinking about micropolicies.} and a partial policy
step function that
refines the machine step function by allowing or disallowing a step to be taken
based on its auxiliary state. This model will be made more precise in
\cref{sec:prelim}. It is flexible enough to capture any kind of mechanism used
for dynamic enforcement of properties of interest, including the following
representative families of policies.

@@

\paragraph*{Null policy}
%
The simplest possible policy is the neutral policy that carries no information
and does nothing. Its set of policy states is the unit type, and its policy step
function is the constant unit function, thereby allowing all calls to the machine
step function. Its initialization function is defined on all states. This
corresponds to the lifting of our machine model to the policy setting. Machines
that fulfill a property by construction will also fulfill that property with the
null policy.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\bcp{Hey, we should have an abstract \ldots}
\end{abstract}

@@

%\apt{Need to clarify to what extent our properties depend on the
%particular compilation style, calling sequences, etc. illustrated here.}

\section{Threat Model and Assumptions}
\label{sec:threat}

@@

Stack safety is meant to protect a caller from a confused or malicious callee's
attempts to access private data or hijack its control flow. Our
attacker model reflects this intention by assuming that the attacker
can run arbitrary machine code to try to access or overwrite its caller's stack data
or disrupt its control state.  Hardware and timing attacks are out of
scope.
%\leo{Anything else to add here?}
%\leo{Maybe a good place to discuss calling conventions and that they
%  are essentially enforced by the policy or sth?}
%\rb{Intuition that the callee can do anything in principle, but the machine will
%halt if it does not follow the calling convention, etc.?}

We assume we are given a machine language program, together with certain auxiliary declarations
about its structure: the code footprint of each function and the locations of instructions
representing calls, returns, and function entry points.
Although this information would typically come from a compiler, we do not
assume anything about the provenance of the machine
code. We assume that the machine has a program counter $\PCname$
and stack pointer $\rsp$ and can perform the usual
loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.

To make our stack safety definitions precise, we make some simple assumptions about
memory layout; our definitions could easily be changed to be parametric over
these assumptions, at the cost of slightly greater complexity.  Memory 
consists of an instruction region and a stack that grows upwards; for simplicity,
we omit a heap. (Integrating a memory safety property for the heap in the style of
\citet{AzevedoDeAmorim17} is a natural extension that we discuss
in~\Cref{sec:future}.)  The stack has a conventional structure, with
contiguously allocated per-function frames.
The boundary between the frames of a caller and callee is given by the $\rsp$ register
at the time of the call.  The stack can be used to
pass arguments or return results in the usual way, by placing them
at the top of the caller's frame. Functions can also communicate via registers; indeed, our 
definitions say nothing at all about registers (except $\PCname$ and $\rsp$).

Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
which are not fully explicit in machine code.  We define a call to be
any instruction declared as such in the program.
If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
of well-bracketed control flow relates this definition of return to the program's
declared set of return points.)
Our property definitions make no other detailed assumptions about call or return sequences, although 
some of the enforcement mechanisms we consider do.

Enforcement mechanisms are modeled as {\em policies} that add some auxiliary
policy state to the ordinary machine state\bcp{This may be surprising to
people who are not thinking about micropolicies.\apt{added para below}} and a partial policy
step function that 
refines the machine step function by allowing or disallowing a step to be taken
based on its auxiliary state.   Enforcement works by halting the program before
it can perform an action that would violate the security property; our properties
are naturally therefore \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

We model ordinary program state and policy enforcement state separately to make it 
easier to use well-established concepts from language-based security.  
In particular, confidentiality has an elegant extensional characterization in terms of
\emph{noninterference}: a program preserves the confidentiality of certain data if
\emph{varying} that data would not change the program's observable behavior~\citep{??}.
In applying this idea to a system that incorporates policy enforcement, only
ordinary program data should be varied, whereas any associated policy enforcement
state should be left alone; our model makes this distinction obvious.
%That is, we compare the actual execution of the program with a \emph{hypothetical}
%execution where data has been varied. 
We also introduce a notion of observational integrity that relies on the idea of
\emph{rolling back} program state; again, our model makes it easy to describe
rollback of ordinary program data while leaving policy enforcement data unchanged.

@@

This enforcement model is flexible enough to capture any kind of mechanism used
for dynamic enforcement of properties of interest, including the following
representative families of policies.

\apt{rest of this section still needs rethink}
\paragraph*{Null policy}
%
The simplest possible policy is the neutral policy that carries no information
and does nothing. Its set of policy states is the unit type, and its policy step
function is the constant unit function, thereby allowing all calls to the machine
step function. Its initialization function is defined on all states. This
corresponds to the lifting of our machine model to the policy setting. Machines
that fulfill a property by construction will also fulfill that property with the
null policy.
